Exploratory Data Analysis: Kickstarter Projects 
========================================================
#### by Ryan Gee

```{r  echo=FALSE, global_options}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
#install.packages("tidyverse")
library(ggplot2)
library(tidyverse)
library(lubridate)
library(knitr)
#library(kableExtra)
```

```{r echo=FALSE, Load_the_Data}
ks <- read.csv('data/ks-projects-201801.csv', fileEncoding='latin1')
```

# The Dataset and Pre-Processing
Kickstarter is a crowdfunding platform that allows a global community to back 
and fund people's creative projects and ideas. Kickstarter was launched in 2009.
In 2018, 15 billion people have pledged $4 Billion dollars, and 150 thousand 
projects have been successfully funded.

This dataset was downloaded from Kaggle
(https://www.kaggle.com/kemical/kickstarter-projects). According to Kaggle this
data was consumed in January 2018.


```{r echo=FALSE, Univariate_Plots}
sprintf('There are %s rows and %s columns in this Kickstarter Data set.',
        dim(ks)[1], dim(ks)[2])
```

The columns and their data type are shown below:

```{r echo=FALSE}
sapply(ks, class)
```

Here are the top 3 rows of the Kickstarter Dataset:

```{r echo=FALSE, results = 'asis'}
kable(head(ks,3), caption = 'Head of Kickstarter Dataset')
```

It looks like there are two similar columns: one is usd.pledged and the other
is usd_pledged_real. After taking a look at the data overview tab on Kaggle,
I have decided to use only the usd.pledged.real column since this column was
converted using the more accurate Fixer.io API. Similarly, I will be using
usd_goal_real instead of goal. Finally, I will drop the currency column since
everything has been converted to USD.

```{r echo=FALSE}
ks <- ks[,!(names(ks) %in% c('usd.pledged', 'goal', 'currency'))]
```

Next, I need to check if there are any NaNs in the dataset:
```{r}
sprintf('There are not any NaNs in the dataset: %s', !any(is.na(ks)))
```

To make life easier in the future, I will convert the deadline and launched
columns to dates and datetimes using the lubridate library:

```{r}
ks$launched <- ymd_hms(ks$launched)
ks$deadline <- ymd(ks$deadline)
sprintf('Launched and Deadlines have been converted to Dates: %s',
        sapply(ks$deadline[1], is.Date))
```

Here is the first few rows of the cleaned up dataset:

```{r results = 'asis'}
kable(head(ks,3), caption = 'Cleaned Kickstarter Dataset')
```


# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

To get a feel of what the Kickstarter dataset looks like, I'll create some quick
Histograms that count the number of projects for some of our categorical data.

I'll first take a look at the **number of projects per Main Category**.

```{r}
sprintf('There are %s different Main Categories.', nlevels(ks$main_category))
ggplot(aes(x=main_category), data = ks)+
  geom_histogram(stat='count') +
  ggtitle('Number of Projects by Main Categories') +
  theme(axis.text.x=element_text(size=8, angle=90)) +
  xlab('')
```

Film and Video projects are clearly the most popular main category, while
Dance is the least popular. Let's take a look at the projects by sub-category.

```{r}
sprintf('There are %s different Sub Categories.', nlevels(ks$category))
ggplot(aes(x=category), data = ks)+
  geom_histogram(stat='count') +
  ggtitle('Number of Projects by Sub Categories') +
  theme(axis.text.x=element_text(size=8, angle=90)) +
  xlab('')
```

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!